---
title: "Final_Project_1"
author: "Dane Hankamer"
date: "7/29/2019"
output: pdf_document
---

Read in CSV files and combine the data (use all 5 studies!)
```{r}
library(plyr)
library(dplyr)
library(ggplot2)
library(data.table)
library(sm)
A = read.csv("Study_A.csv", stringsAsFactors=TRUE)
B = read.csv("Study_B.csv", stringsAsFactors=TRUE)
C = read.csv("Study_C.csv", stringsAsFactors=TRUE)
D = read.csv("Study_D.csv", stringsAsFactors=TRUE)
E = read.csv("Study_E.csv", stringsAsFactors=TRUE)
total_data = rbind(A, B, C, D)
# total_data = rbind(total_data[, names(E)], E)
# str(total_data)
# summary(total_data)
# cor(total_data[sapply(total_data, is.numeric)])
```

Separate dataframes based on control versus treatment patients
```{r}
c_data = filter(total_data, total_data$TxGroup == "Control")
t_data = filter(total_data, total_data$TxGroup == "Treatment")
c_data = data.frame(c_data$PatientID, c_data$VisitDay, c_data$PANSS_Total)
t_data = data.frame(t_data$PatientID, t_data$VisitDay, t_data$PANSS_Total)
```

Check average PANSS scores for control and treatment data
```{r}
c_pans_avg = summarize(c_data, mean(c_data$PANSS_Total))
t_pans_avg = summarize(t_data, mean(t_data$PANSS_Total))
# c_pans
# t_pans
```

1) Group by unique patient ID for the control dataframe (already done)
2) Get differences in PANSS scores per each new visit
3) Calculate average PANSS score change per visit for each patient by taking sum of differences over PatientID counts, where our average includes a difference of 0 for the first visit (some patients only showed up once so don't want to do Sum(Diff)/(Count-1) to get true average PANSS score change and divide by 0)
```{r}
panss_diff = ave(c_data$c_data.PANSS_Total, c_data$c_data.PatientID, FUN=function(x) c(0, diff(x)))
c_data["c_data.panss_diff"] = panss_diff
c_data = c_data[,c(1,4)]
```
```{r}
control_panss_avgs = aggregate(c_data$c_data.panss_diff~c_data$c_data.PatientID, c_data, mean)
# control_panss_avgs
```

Repeat for treatment group, plot overlapping histograms
```{r}
panss_diff_t = ave(t_data$t_data.PANSS_Total, t_data$t_data.PatientID, FUN=function(x) c(0, diff(x)))
t_data["t_data.panss_diff"] = panss_diff_t
t_data = t_data[,c(1,4)]
```
```{r}
treatment_panss_avgs = aggregate(t_data$t_data.panss_diff~t_data$t_data.PatientID, t_data, mean)
hist(control_panss_avgs$`c_data$c_data.panss_diff`,
     main="Histogram for Avg. Change in PANSS Score",
     xlim=c(-30,30),
     ylim=c(0,600),
     breaks=25, 
     col=rgb(0,0,1,0.8),
     xlab="Avg. PANSS Score Change / Visit / PatientID")
hist(treatment_panss_avgs$`t_data$t_data.panss_diff`,
     col=rgb(1,0,0,0.8),
     add=T)
legend("topright", c("Control", "Treatment"), fill=c(rgb(0,0,1,0.8), rgb(1,0,0,0.8)))
box()
```

```{r}
control_density = data.frame(x=control_panss_avgs$`c_data$c_data.panss_diff`, vec="Control")
treatment_density = data.frame(x=treatment_panss_avgs$`t_data$t_data.panss_diff`, vec="Treatment")
df = rbind(control_density, treatment_density)

gg = ggplot(data=df, aes(x=x,fill=vec)) + geom_density(alpha=0.2)
gg + labs(title="Density Plot for Average Change in PANSS Score", x="Avg. PANSS Score Change")
```

```{r}
cat("Mean score change for control patients/visit: ", mean(control_panss_avgs$`c_data$c_data.panss_diff`), "\n")
cat("PANSS score std dev for control patients/visit: ", sd(control_panss_avgs$`c_data$c_data.panss_diff`), "\n")
cat("Mean score change for treatment patients/visit: ", mean(treatment_panss_avgs$`t_data$t_data.panss_diff`), "\n")
cat("PANSS score std dev for treatment patients/visit: ", sd(treatment_panss_avgs$`t_data$t_data.panss_diff`), "\n")
```

Run variance test (Fisher's F-test) to see if the data passes the homoscedasticity test (see if variances are homogenous)
```{r}
control = control_panss_avgs$`c_data$c_data.panss_diff`
treatment = treatment_panss_avgs$`t_data$t_data.panss_diff`
var.test(control, treatment)
```

Fisher's F-test yields a p-value > 0.05, so we will assume that the samples (control versus treatment) are homogenous. Now we run a two-sample t-test with var.equal=TRUE to see if the treatment has an effect on schizophrenia.
```{r}
treatment_effect_test = t.test(control, treatment, var.equal=TRUE)
treatment_effect_test
```