---
title: "Final_Project_4.Rmd"
author: "Dane Hankamer"
date: "8/7/2019"
output:
  pdf_document: default
  html_document: default
---

Conclusions: We see extreme differences in the percent of scores flagged/assigned to CS based on the test country. We will use this to our advantage in Problem 4. There does not seem to be a significant difference between the averages for Total_PANSS_Score between those tests that were flagged/assigned to CS and those that passed. Also, there are vast differences between the percent of flagged/assigned to CS tests based on the study (A,B,C,D).
```{r}
library(plyr)
library(dplyr)
library(ggplot2)
library(data.table)
library(sm)
A = read.csv("Study_A.csv", stringsAsFactors=TRUE)
B = read.csv("Study_B.csv", stringsAsFactors=TRUE)
C = read.csv("Study_C.csv", stringsAsFactors=TRUE)
D = read.csv("Study_D.csv", stringsAsFactors=TRUE)
E = read.csv("Study_E.csv", stringsAsFactors=TRUE)
total_data = rbind(A, B, C, D)
flagged_data = subset(total_data, total_data$LeadStatus!="Passed")
passed_data = subset(total_data, total_data$LeadStatus=="Passed")
# summary(flagged_data)
```

```{r}
# summary(passed_data)
```

```{r}
options(max.print=100000)
n_passed_data = round(total_data[sapply(total_data, is.numeric)], 2)
n_flagged_data = round(total_data[sapply(total_data, is.numeric)], 2)
passed_mat = as.matrix(n_passed_data)
flagged_mat = as.matrix(n_flagged_data)
# cor(passed_mat, flagged_mat)
```

```{r}
# summary(E)
```

```{r}
table(total_data$Country)
count(total_data)
```
